---
title: "ICIJ IMDD 20181219 Data Overview"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
  html_notebook:
    toc: yes
    toc_depth: 2
---


```{r, echo=F}
UsePackage <- function(p) {
  if (!is.element(p, installed.packages()[,1]))
    install.packages(p, dep = TRUE)
  require(p, character.only = TRUE)
}
UsePackage("dplyr")
UsePackage("knitr")
UsePackage("tidyr")
UsePackage("lubridate")
UsePackage("ggplot2")
# Loads
load(file="../data/interim/imdd.Rdata")
device=imdd$device
manuf=imdd$manuf
event=imdd$event
rm(imdd)
# Fix date
event$date_initiated_by_firm <- as.Date(event$date_initiated_by_firm)
event$date_posted <- as.Date(event$date_posted)
event$date <- as.Date(event$date, format="%Y-%m-%d")
event$create_date <- as.Date(event$create_date, format="%Y-%m-%d")
event$new_date <- ifelse(!is.na(event$date_initiated_by_firm), event$date_initiated_by_firm, ifelse(!is.na(event$date_posted), event$date_posted, ifelse(!is.na(event$create_date), event$create_date, event$date)))
event$new_date <- as.Date(event$new_date, origin="1970-01-01")
# Safety only
safety <- event %>% filter(type != "Recall")
```

Three datasets, device, event, manuf. Dimensions are...

```{r, echo=F}
print(dim(device))
print(dim(event))
print(dim(manuf))
```

## Univars

### Manufacturers

#### Names

Unique Names: `r length(unique(manuf$name))`

Missing Names: `r sum(is.na(manuf$name) | manuf$name == "")`

Top Names:

```{r, echo=F}
kable(top_n(arrange(dplyr::count(manuf, name), desc(n)), 10))
```

#### Parents

Unique Parents: `r length(unique(manuf$parent_company))`

Missing Parents: `r sum(is.na(manuf$parent_company) | manuf$parent_company == "")`

Top Parents:

```{r, echo=F}
kable(top_n(arrange(dplyr::count(manuf, parent_company), desc(n)), 10))
```

### Devices

#### Description

Unique Description: `r length(unique(device$description))`

Missing Description: `r sum(is.na(device$description) | device$description == "")`

Top Description:

```{r, echo=F}
kable(top_n(arrange(dplyr::count(device, description), desc(n)), 10))
```

#### Classifications

Unique Classifications: `r length(unique(device$classification))`

Missing Classifications: `r sum(is.na(device$classification) | device$classification == "")`

Top Classifications:

```{r, echo=F}
kable(top_n(arrange(dplyr::count(device, classification), desc(n)), 10))
```

#### Risk Classes

Unique Risk Classes: `r length(unique(device$risk_class))`

Missing Risk Classes: `r sum(is.na(device$risk_class) | device$risk_class == "")`

Top Risk Classes:

```{r, echo=F}
kable(top_n(arrange(dplyr::count(device, risk_class), desc(n)), 10))
```

#### Countries

Unique Countries: `r length(unique(device$country))`

Missing Countries: `r sum(is.na(device$country) | device$country == "")`

Top Countries:

```{r, echo=F}
kable(top_n(arrange(dplyr::count(device, country), desc(n)), 10))
```

#### Date Firm Initiation

Missing dates: `r sum(is.na(event$date_initiated_by_firm))`

```{r, echo=F}
kable(table(year(event$date_initiated_by_firm)))
```

#### Date Posted

Missing dates: `r sum(is.na(event$date_posted))`

```{r, echo=F}
kable(table(year(event$date_posted)))
```

#### Create Date

Missing dates: `r sum(is.na(event$create_date))`

```{r, echo=F}
kable(table(year(event$create_date)))
```

#### Date (not otherwise specified)

Missing dates: `r sum(is.na(event$date))`

```{r, echo=F}
kable(table(year(event$date)))
```

#### Any Date

Missing all dates: `r sum(is.na(event$new_date))`

```{r, echo=F}
kable(table(year(event$new_date)))
```

#### Year by Country

```{r, echo=F}
yrcts <- event %>% mutate(year=year(as.Date(new_date, origin="1970-01-01"))) %>%
  filter(country %in% c("CHE", "ESP", "FIN", "FRA", "NLD")) %>%
  count(year, country) %>% rename(Country=country)
ggplot(data=yrcts, aes(x=year, y=n, color=Country)) +
    geom_line(size=1.5) + theme_minimal(base_size=16) +
  xlab("Year of Event") + ylab("Number of Events") + ggtitle("Events by EU Country")
pops <- data.frame(Country=c("CHE", "ESP", "FIN", "FRA", "NLD"),
                   Population=c(8.526, 46.733, 5.522, 67.372, 17.284))
yrcts <- merge(yrcts, pops, all.x=T) %>% mutate(n_adj=n / Population)
ggplot(data=yrcts, aes(x=year, y=n_adj, color=Country)) +
    geom_line(size=1.5) + theme_minimal(base_size=16) +
  xlab("Year of Event") + ylab("Events / Population") + ggtitle("Events by EU Country - Population Adjusted")
```


#### Action Classification

Unique Classification: `r length(unique(event$action_classification))`

Missing Classification: `r sum(is.na(event$action_classification) | event$action_classification == "")`

Top Classification:

```{r, echo=F}
kable(top_n(arrange(dplyr::count(event, action_classification), desc(n)), 10))
```

#### Action Classification by Country

```{r, echo=F}
acbyc <- event %>% count(country, action_classification) %>% spread(action_classification, n) %>% rename(Missing=V1)
kable(acbyc)
```

#### Country

Unique Country: `r length(unique(event$country))`

Missing Country: `r sum(is.na(event$country) | event$country == "")`

Top Country:

```{r, echo=F}
kable(top_n(arrange(dplyr::count(event, country), desc(n)), 13))
```

#### Reason

Unique Reason: `r length(unique(event$reason))`

Missing Reason: `r sum(is.na(event$reason) | event$reason == "")`

Top Reason:

```{r, echo=F}
kable(top_n(arrange(dplyr::count(event, reason), desc(n)), 25))
```

#### Determined Cause

Unique Cause: `r length(unique(event$determined_cause))`

Missing Cause: `r sum(is.na(event$determined_cause) | event$determined_cause == "")`

Top Cause:

```{r, echo=F}
kable(top_n(arrange(dplyr::count(event, determined_cause), desc(n)), 10))
```

#### Action Summary

Unique Action: `r length(unique(event$action_summary))`

Missing Action: `r sum(is.na(event$action_summary) | event$action_summary == "")`

Top Action:

```{r, echo=F}
kable(top_n(arrange(dplyr::count(event, action_summary), desc(n)), 25))
```

#### Type

Unique Type: `r length(unique(event$type))`

Missing Type: `r sum(is.na(event$type) | event$type == "")`

Top Type:

```{r, echo=F}
kable(top_n(arrange(dplyr::count(event, type), desc(n)), 10))
```

### For Safety Reasons

#### Safety Reason

Unique Reason: `r length(unique(safety$reason))`

Missing Reason: `r sum(is.na(safety$reason) | safety$reason == "")`

Top Reason:

```{r, echo=F}
kable(top_n(arrange(dplyr::count(safety, reason), desc(n)), 25))
```

#### Safety Determined Cause

Unique Cause: `r length(unique(safety$determined_cause))`

Missing Cause: `r sum(is.na(safety$determined_cause) | safety$determined_cause == "")`

Top Cause:

```{r, echo=F}
kable(top_n(arrange(dplyr::count(safety, determined_cause), desc(n)), 10))
```

#### Safety Action Summary

Unique Action: `r length(unique(safety$action_summary))`

Missing Action: `r sum(is.na(safety$action_summary) | safety$action_summary == "")`

Top Action:

```{r, echo=F}
kable(top_n(arrange(dplyr::count(safety, action_summary), desc(n)), 25))
```

*Disclaimer*

This notebook represents exploratory analysis. It has not been validated. It should not be used to inform any decisions, nor should interpretations from this analysis be made regarding the IMDD.